### 2025-06-09 Abstract.tex
add Abstract for the research paper

\documentclass[conference]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{cite}
\usepackage{url}
\usepackage[none]{hyphenat}

\title{Anti-Plagiarism System for Exam Monitoring}

\author{
    \IEEEauthorblockN{Valentin Pletea-Marinescu}
    \IEEEauthorblockA{
        \textit{National University of Science and Technology POLITEHNICA Bucharest}\\
        Email: \texttt{pletea.valentin2003@gmail.com}
    }
}

\begin{document}

\maketitle

\begin{abstract}
Academic integrity represents a fundamental challenge in modern education systems, with plagiarism rates increasing globally across all educational levels. Traditional exam monitoring approaches rely primarily on screen surveillance and human oversight, creating significant vulnerabilities in detecting sophisticated cheating behaviors during online and remote examinations.

This research develops a comprehensive anti-plagiarism monitoring system using advanced machine learning and computer vision technologies to address these limitations. The system integrates sophisticated gaze tracking algorithms with YOLO-based object detection models through a modular software architecture that combines facial landmark detection with real-time behavioral analysis. Custom algorithms continuously analyze candidate gaze patterns while convolutional neural networks detect unauthorized objects and suspicious materials, processing video streams efficiently using optimized OpenCV libraries.

Experimental testing demonstrates highly effective operation on standard CPU-based systems, successfully identifying suspicious gaze patterns, unauthorized devices, and anomalous behaviors while maintaining consistent real-time performance. The system achieved robust detection capabilities during controlled testing scenarios. This research provides a comprehensive open-source solution for educational institutions, enabling widespread deployment using standard computing hardware without requiring specialized equipment.
\end{abstract}

\begin{IEEEkeywords}
Academic integrity, Educational technology, Computer vision, Gaze tracking, Object detection, Real-time systems, Machine learning, Convolutional neural networks, Image processing
\end{IEEEkeywords}

\section{Introduction}

\end{document}